---
title: "Appendix 4: Generic model code"
output: pdf_document
---


Here we provide generic model code that closely follows the model description described in the main text, without the extra data structures to account for data heterogeneity (as found in Appendix 3).

# Generic model code

```{r, eval = FALSE}
model_code <- NimbleCode({
  
  #################
  #################
  # Process model #
  #################
  #################
  
  #############################
  # Integral projection model #
  #############################
  
  # intra-annual change
  for (y in 1:n_year) {
    for (t in 1:(Tmax-1)) {
      
      # Equation 1
      ## project
      N[t + 1, y, 1:n_size] <- kernel[1:n_size, 1:n_size, t] %*%
        (N[t, y, 1:n_size] - C_T[t, y, 1:n_size]) +
        recruit_intro[t, y] * R[y, 1:n_size] # introduce recruits, t = 6

      # Equation 2
      ## kernel
      kernel[1:n_size, 1:n_size, t] <- seasonal_growth(
        xinf, gk, sigma_G, A, ts, D[t, 1:2], n_size, pi, x[1:n_size], 
        lower[1:n_size], upper[1:n_size]) * S[t, y, 1:n_size]
      # Equation 7
      ## natural survival
      S[t, y, 1:n_size] <- survival(alpha, beta[y], D[t, 1:2], x[1:n_size])
    }
  }
  
  # Equation 8
  ## year-specific natural mortality (process error, includes mark-recapture)
  for (y in 1:(n_year + 1)) {
    beta[y] ~ dgamma(beta_alpha, beta_theta)
  }
  
  ## inter-annual change
  for (y in 1:(n_year - 1)) {
    for (k in 1:n_size) {
      
      # Equation 9-10
      ## overwinter survival and seasonal growth
      ## (use stochastic N node)
      N_stoch[y, k] ~ dbinom(size = seasonal_growth(xinf, gk, sigma_G, A, ts, 
                                                    D[Tmax, 1:2], k, pi, y[k], 
                                                    lower[k], upper[k]) %*% 
                               (N[Tmax, y, k] - C_T[Tmax, y, k]),
                             prob = S_o[y, k)
      
    }
    
    # fill deterministic N node with stochastic N node
    N[1, y + 1, 1:n_size] <- N_stoch[y, 1:n_size]
    
    # Equation 11
    ## size- and density-dependent overwinter survival
    S_o[y, 1:n_size] <- overwinter_survival(alpha_o[y], 
                                            sum(N[Tmax, y, 1:n_size]), 
                                            x[1:n_size])
    
    # Equation 12
    ## year-specific intensity of overwinter mortality
    alpha_o[y] ~ dgamma(alpha_o_alpha, alpha_o_theta)
  }
  
  #####################################################
  # Initial population density and annual recruitment #
  #####################################################
  
  # Equation 12
  ## get initial size-structured abundance of adults - year 1
  N[1, 1, 1:n_size] <- get_init_adult(mu_A, sigma_A, lower[1:n_size], 
                                      upper[1:n_size], lambda_A)
  
  # Equation 14
  ## annual abundance of recruits
  for (m in 1:n_year) {
   lambda_R[m] ~ T(dnorm(mu_lambda, sd = sigma_lambda), 0, Inf)
  }
  
  # Equation 15
  ## annual size distribution of recruits
  R[1:n_year, 1:n_size] <- get_init_recruits(mu_R, sigma_R, lower[1:n_size], 
                                             upper[1:n_size], 
                                             lambda_R[1:n_year], n_year,
                                             n_size)
  
  #####################
  #####################
  # Observation model #
  #####################
  #####################
  
  for (y in 1:n_year) {
    for (t in 1:Tmax) {
      for (k in 1:n_size) {
        
        # Equation 16
        ## binomial distribution with total crabs removed, n_R
        C_T[t, y, k] ~ dbinom(size = round(N[t, y, k]), prob = p[t, y, k])

        # Equation 17
        ## dirichlet-multinomial mixture, conditional probability of capture
        alpha_D[t, 1:totalo, y, k] <- p_C[t, 1:totalo, y, k] * n_p_dir
        C[t, 1:totalo, y, k] ~ ddirchmulti(alpha = alpha_D[t, 1:totalo, y, k],
                                             size = C_T[t, y, k])
      }
      
      # Equations 18 - 20
      ## calculate hazard rate
      hazard[t, 1:totalo, y, 1:n_size] <- calc_hazard(
        totalo, n_size, h_F_max, h_F_k, h_F_0, h_S_max, h_S_k, h_S_0, h_M_max,
        h_M_A, h_M_sigma, f_index[t, 1:totalo, y], s_index[t, 1:totalo, y],
        m_index[t, 1:totalo, y], soak_days[t, 1:totalo, y], x[1:n_size]
        )

      # Equation 21
      ## total capture probability
      p[t, y, 1:n_size] <- calc_prob(totalo, n_size,
                                     hazard[t, 1:totalo, y, 1:n_size])
      
      # mean conditional probability of capture
      p_C[t, 1:totalo, y, 1:n_size] <- calc_cond_prob(totalo, n_size,
                                                      hazard[t, 1:totalo,
                                                             y, 1:n_size])
      
    }
  }
  
  
  ###############################
  ###############################
  # Integrated population model #
  ###############################
  ###############################
  
  #######################
  # Mark-recapture data #
  #######################
  
  # Equation 24
  ## draw recaptured samples, m2, from projected marked samples, n1_project
  for (k in 1:n_size) {
    m2[k] ~ dbinom(size = round(n1_project[k]), prob = p_mc[k])
  }
  
  # Equation 25
  ## total capture probability with mark-recapture data
  p_mc[1:n_size] <- calc_prob(totalo_mc, n_size,
                              hazard_mc[1:totalo_mc, 1:n_size])
  
  # size-dependent hazard rates with mark-recapture data
  hazard_mc[1:totalo_mc, 1:n_size] <- calc_hazard(
    totalo_mc, n_size, h_F_max, h_F_k, h_F_0, h_S_max, h_S_k, h_S_0, h_M_max,
        h_M_A, h_M_sigma, f_index_mc[1:totalo_mc], s_index_mc[1:totalo_mc,],
        m_index_mc[1:totalo_mc], soak_days[1:totalo_mc], x[1:n_size]
  )
  
  # Equation 26
  ## apply kernel from time of marking to time of recapture
  n1_project[1:n_size] <- kernel_mc[1:n_size, 1:n_size] %*% n1[1:n_size]
  
  kernel_mc[1:n_size, 1:n_size] <- seasonal_growth(xinf, gk, sigma_G, A, ts, 
                                                   D_mc[1:2], n_size,  pi, 
                                                   x[1:n_size], lower[1:n_size], 
                                                   upper[1:n_size]) * 
    survival(alpha, beta[5], D_mc[1:2], x[1:n_size])
  
  ################
  # Growth model #
  ################
  
  # incorporated hierarchically as informative priors in the unified model
  # asymptotic size
  xinf ~ dnorm(inf_prior[1, 1], sd = inf_prior[1, 2])
  # growth rate
  gk ~ dnorm(inf_prior[2, 1], sd = inf_prior[2, 2])
  # amplitude of growth oscillations
  A ~ dnorm(inf_prior[3, 1], sd = inf_prior[3, 2])
  # inflection point of growth oscillations
  ts ~ dnorm(inf_prior[4, 1], sd = inf_prior[4, 2])
  
  
  
  
  ################
  ################
  # Vague priors #
  ################
  ################
  
  ##
  # size selectivity parameters
  ##

  # minnow max. hazard rate
  h_M_max ~ dunif(priors[1, 1], priors[1, 2])
  # minnow max. size of capture
  h_M_A ~ dunif(priors[2, 1], priors[2, 2])
  # minnow sigma of gaussian size selectivity curve
  h_M_sigma ~ dunif(priors[3, 1], priors[3, 2])
  # fukui max. hazard rate
  h_F_max ~ dunif(priors[4, 1], priors[4, 2])
  # fukui k of logistic size selectivity curve
  h_F_k ~ dunif(priors[5, 1], priors[5, 2])
  # fukui midpoint of logistic size selectivity curve
  h_F_0 ~ dunif(priors[6, 1], priors[6, 2])
  # shrimp max. hazard rate
  h_S_max ~ dunif(priors[7, 1], priors[7, 2])
  # shrimp k of logistic size selectivity curve
  h_S_k ~ dunif(priors[8, 1], priors[8, 2])
  # shrimp midpoint of logistic size selectivity curve
  h_S_0 ~ dunif(priors[9, 1], priors[9, 2])

  ##
  # IPM - natural mortality
  ##

  # gamma distributions shape for instantaneous intensity of mortality
  beta_alpha ~ dunif(priors[10, 1], priors[10, 2])
  # gamma distributions rate for instantaneous intensity of mortality
  beta_theta ~ dunif(priors[11, 1], priors[11, 2])
  # size-dependent overwinter natural mortality, shared across all sites
  alpha ~ dunif(priors[12, 1], priors[12, 2])
  # gamma distribution shape - instantaneous intensity of overwinter mortality
  alpha_o_alpha ~ dunif(priors[13, 1], priors[13, 2])
  # gamma distribution rate - instantaneous intensity of overwinter mortality
  alpha_o_theta ~ dunif(priors[14, 1], priors[14, 2])

  ##
  # IPM - growth
  ##
  
  # growth error
  sigma_G ~ dunif(priors[15, 1], priors[15, 2])

  ##
  # observation process
  ##
  # dirichlet multinomial (overdispersion in count data)
  ro_dir ~ dbeta(priors[16, 1], priors[16, 2])
  n_p_dir <- (1 - ro_dir) / ro_dir

  ##
  # initial population density and annual recruitment
  ##

  # initial adult size (mean and sd)
  mu_A ~ dunif(priors[17, 1], priors[17, 2])
  sigma_A ~ dunif(priors[18, 1], priors[18, 2])

  # initial recruit size (mean and sd)
  mu_R ~ dunif(priors[19, 1], priors[19, 2])
  sigma_R ~ dunif(priors[20, 1], priors[20, 2])

  # abundance of recruits (mean and sd)
  mu_lambda ~ dunif(priors[21, 1], priors[21, 2])
  sigma_lambda ~ dunif(priors[22, 1], priors[22, 2])

  # abundance of adults in year 1
  lambda_A ~ dunif(priors[23, 1], priors[23, 2])
  
  
})
```

# Nimble functions

## Process model functions

### Initial population density of adults

This function corresponds to equation 1 in main text.

```{r, eval = FALSE}
# initial size distribution of adults - year 1
get_init_adult <- nimbleFunction (
  
  run = function(mu_A = double(0), sigma_A = double(0),
                 lower = double(1), upper = double(1),
                 lambda_A = double(0))
  {
    returnType(double(1))
    
    # moment match from normal to gamma
    var <- sigma_A ^ 2
    shape <- mu_A ^ 2 / var
    rate <- mu_A / var
    prop_adult <- pgamma(q = upper, shape = shape, rate = rate) -
      pgamma(q = lower, shape = shape, rate = rate)
    
    # get initial size-structured abundance of adults
    out <- prop_adult * lambda_A

    return(out)
  }
)
```

### Size-structured recruitment

This function corresponds to equation 15 in main text.

```{r, eval = FALSE}
# annual size distributions of recruit
get_init_recruits <- nimbleFunction (
    
  run = function(mu_R = double(0), sigma_R = double(0),
                 lower = double(1), upper = double(1),
                 lambda_R = double(1), n_year = double(0),
                 n_size = double(0))
  {
    returnType(double(2))
    
    # create empty array
    out <- matrix(NA, ncol = n_size, nrow = n_year)

    var <- sigma_R ^ 2
    shape <- mu_R ^ 2 / var
    rate <- mu_R / var
    prop_recruit <- pgamma(q = upper, shape = shape, rate = rate) -
      pgamma(q = lower, shape = shape, rate = rate)
      
    # get initial size-structured abundance of recruits
    for (i in 1:n_year) {
      out[i, ] <- prop_recruit[1:n_size] * lambda_R[i]
    }

    return(out)
  }
)
```

### Seasonal growth

This function corresponds to equations 3-5 in main text.

```{r, eval = FALSE}
# function for seasonal growth kernel
seasonal_growth <- nimbleFunction (
  run = function(xinf = double(0), k = double(0),
                 sigma_G = double(0), A = double(0),
                 ts = double(0), D = double(1),
                 n_size = double(0), pi = double(0), x = double(1),
                 lower = double(1), upper = double(1))
  {
    returnType(double(2))

    # create empty array
    array <- matrix(NA, ncol = n_size, nrow = n_size)

    # season adjusted params
    S_t <- (A * k / (2 * pi)) * sin(2 * pi * (D[2] - (1 + ts)))
    S_t0 <- (A * k / (2 * pi)) * sin(2 * pi * (D[1] - (1 + ts)))

    # p(y"|y)
    for (i in 1:n_size) {
      increment <- (xinf - x[i]) *
        (1 - exp(-k * (D[2] - D[1]) - S_t + S_t0))
      mean <- x[i] + increment
      array[1:n_size, i] <- (pnorm(upper, mean, sd = sigma_G) -
                               pnorm(lower, mean, sd = sigma_G))
    }

    # normalize
    for (i in 1:n_size) {
      array[, i] <- array[, i] / sum(array[, i])
    }
  return(array)
}
)
```

### Natural survival

This function corresponds to equation 7 in main text.

```{r, eval = FALSE}
# natural (non-winter) survival
survival <- nimbleFunction (
  
  run = function(alpha = double(0), beta = double(0), D = double(1),
                 x = double(1))
  {
    returnType(double(1))
    
    # get survival rate
    out <- exp(-(D[2] - D[1]) * (beta + alpha / x ^ 2))

    return(out)
  }
)
```

### Density-dependent overwinter survival

This function corresponds to equation 9-10 in main text.

```{r, eval = FALSE}
# density- and size-dependent overwinter survival
overwinter_survival <- nimbleFunction (
  
  run = function(alpha_o = double(0), N = double(1),
                 x = double(1))
  {
    returnType(double(1))
    
    # get probability of survival
    out <- exp(-(alpha_o * sum(N) / x ^ 2))

    return(out)
  }
)
```

## Observation model functions

### Conditional multinomial observation model

This function corresponds to equations 18-20 in main text.

```{r, eval = FALSE}
# calculate trap hazard rate of obs j, based on trap type and soak days
calc_hazard <- nimbleFunction (
  
  run = function(nobs = double(0), n_size = double(0), h_F_max = double(0),
                 h_F_k = double(0), h_F_0 = double(0), h_S_max = double(0),
                 h_S_k = double(0), h_S_0 = double(0), h_M_max = double(0),
                 h_M_A = double(0), h_M_sigma = double(0), f_index = double(1), 
                 s_index = double(1), m_index = double(1), 
                 soak_days = double(1), x = double(1))
  {
    returnType(double(2))

    # create empty array
    array <- array(init = FALSE, dim = c(nobs, n_size))
    
    # loop through observations
    for (j in 1:nobs) {
      # P(capture)
      array[j, 1:n_size] <- size_sel_log(pmax = h_F_max, k = h_F_k,
                                         midpoint = h_F_0, x = x) *
        f_index[j] * soak_days[j] +
        size_sel_log(pmax = h_S_max, k = h_S_k, midpoint = h_S_0, x = x) *
        s_index[j] * soak_days[j] +
        size_sel_norm(pmax = h_M_max, xmax = h_M_A, sigma = h_M_sigma, x = x) *
        m_index[j] * soak_days[j]
    }
      return(array)
  }
)

```

These functions correspond to equation 21 in main text

```{r, eval = FALSE}
# calculate conditional probability of capture
calc_cond_prob <- nimbleFunction (
  run = function(nobs = double(0), n_size = double(0), hazard = double(2))
  {
    returnType(double(2))
    
    # create empty array
    array <- array(init = FALSE, dim = c(nobs, n_size))
    # loop through sizes
    for (k in 1:n_size) {
      #P(capture in trap j | captured at all)
      array[1:nobs, k] <- hazard[1:nobs, k] / sum(hazard[1:nobs, k])
    }
    return(array)
  }
)

# calculate total capture probability
calc_prob <- nimbleFunction (
    
  run = function(nobs = double(0), n_size = double(0), hazard = double(2))
  {
    returnType(double(1))

    # create empty array
    p <- rep(NA, n_size)

    # loop through sizes
    for (k in 1:n_size) {
      # 1 - exp(-P(captured at all))
      p[k] <- 1 - exp(-sum(hazard[1:nobs, k]))
    }
    return(p)
  }
)
```

### Size-selective hazard rates

These functions correspond to equations 18-20 in main text.

```{r, eval = FALSE}
# bell-shaped size selective hazard rate
size_sel_norm <- nimbleFunction (
  # input and output types
  run = function(pmax = double(0), xmax = double(0), sigma = double(0),
                 x = double(1))
  {
    returnType(double(1))
      vector <- pmax * exp(-(x - xmax) ^ 2 / (2 * sigma ^ 2))

    return(vector)
  }
)

# logistic size selective hazard rate
size_sel_log <- nimbleFunction (
  run = function(pmax = double(0), k = double(0), midpoint = double(0),
                 x = double(1))
  {
    returnType(double(1))
    
    vector <- pmax / (1 + exp(-k * (x - midpoint)))

    return(vector)
  }
)
```

### Dirichlet-multinomial mixture

These functions correspond to equation 17 in main text.

```{r, eval = FALSE}
# define dirichlet multinomial mixture
# pdf
ddirchmulti <- nimbleFunction (
  run = function(x = double(1), alpha = double(1), size = double(0),
                 log = integer(0, default = 0)) {
    returnType(double(0))
    logProb <- lgamma(size + 1) - sum(lgamma(x + 1)) + lgamma(sum(alpha)) -
      sum(lgamma(alpha)) + sum(lgamma(alpha + x)) -
      lgamma(sum(alpha) + size)
    if (log) return(logProb)
    else return(exp(logProb))
  })
# random number generator
rdirchmulti <- nimbleFunction (
  run = function(n = integer(0), alpha = double(1), size = double(0)) {
    returnType(double(1))
    if (n != 1) print("rdirchmulti only allows n = 1; using n = 1.")
    p <- rdirch(1, alpha)
    return(rmulti(1, size = size, prob = p))
  })
```