---
title: "Appendix 5: Posterior summaries and trace plots"
output: pdf_document
---
# Appendix 5
```{r, echo = FALSE, results = "hide", message = FALSE, warning=FALSE}
library(tidyverse)
library(patchwork)
library(MCMCvis)
library(bayestestR)
library(kableExtra)
```

```{r, echo = FALSE, results = "hide"}
# create functions
create_traceplot <- function(out, param, names_df) {
  name <- names_df$real_names[which(names_df$short_names == param)]
  plot <- ggplot() +
    geom_line(aes(x = 1:nrow(out[[1]]),
                  y = out[[1]][, param]),
              color = "pink") +
    geom_line(aes(x = 1:nrow(out[[2]]),
                  y = out[[2]][, param]),
              color = "purple") +
    geom_line(aes(x = 1:nrow(out[[3]]),
                  y = out[[3]][, param]),
              color = "blue") +
    geom_line(aes(x = 1:nrow(out[[4]]),
                  y = out[[4]][, param]),
              color = "red") +
    scale_x_continuous(breaks = c(2000, 4000, 6000)) +
    labs(x = "iteration", y = "value") +
    ggtitle(parse(text = name)) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
  return(plot)
}
```

```{r, echo = FALSE, results = "hide"}
# read in samples
out <- readRDS("../data/posterior_samples/savedsamples_IPM.rds")
```

```{r, echo = FALSE, results = "hide"}
# create df with param names
short_names <- c("log_mu_A", "sigma_A", "mu_R", "sigma_R", "xinf", "gk", 
                 "A", "ds", "sigma_G", "d0", "sigma_w", "sigma_y",
                 "beta", "alpha", "alpha_o", 
                 "eps_y[1]", "eps_y[2]", "eps_y[3]", 
                 "sigma_o", "h_F_max", "h_M_max", "h_S_max", "h_F_k", 
                 "h_S_k", "h_F_0", "h_S_0", "h_M_A", "h_M_sigma", "ro_dir", 
                 "lambda_A", "lambda_R[1]", "lambda_R[2]", "lambda_R[3]",
                 "lambda_R[4]", "mu_lambda", "sigma_lambda")

real_names <- c("mu^A", "sigma^A", "mu^R", "sigma[R]",
                "x[infinity]", "k", "A", "d[s]", "sigma[G]", 
                "d[0]", "sigma[w]", "sigma[y]",
                "beta", "alpha", "alpha[o]", "epsilon[1]", "epsilon[2]",
                "epsilon[3]", "sigma[o]",
                "h[F]^max", "h[M]^max", "h[S]^max", "h[F]^k", "h[S]^k",
                "h[F]^0", "h[S]^0", "h[M]^A", "h[M]^sigma", "rho",
                "lambda^A", "lambda[1]^R", "lambda[2]^R", "lambda[3]^R",
                "lambda[4]^R", "mu^lambda", "sigma^lambda")

latex_names <- c("\\mu^A", "\\sigma^A", "\\mu^R", "\\sigma_R",
                 "x_{\\infty}", "k", "A", "d_s", "\\sigma_G",
                 "d_0", "\\sigma_w", "\\sigma_y",
                 "\\beta", "\\alpha", "\\alpha_o", "\\epsilon_1",
                 "\\epsilon_2", "\\epsilon_3", 
                 "\\sigma_o", "h_F^{max}", "h_M^{max}", "h_S^{max}",
                 "h_F^k", "h_S^k", "h_F^0", "h_S^0", "h_M^A", "h_M^{\\sigma}",
                 "\\rho", "\\lambda^A", "\\lambda_1^R", "\\lambda_2^R",
                 "\\lambda_3^R", "\\lambda_4^R", "\\mu^{\\lambda}",
                 "\\mu^{\\sigma}")

names_df <- as.data.frame(cbind(short_names, real_names, latex_names))
colnames(names_df) <- c("short_names", "real_names", "latex_names")
```

```{r, echo = FALSE, results = "hide"}
# separate parameters into three sets
params1 <- short_names[1:15]
params2 <- short_names[16:30]
params3 <- short_names[31:36]
plots1 <- create_traceplot(out, params1[1], names_df)
plots2 <- create_traceplot(out, params2[1], names_df)
plots3 <- create_traceplot(out, params3[1], names_df)
for (i in 2:length(params1)) {
  new_plot <- create_traceplot(out, params1[i], names_df)
  plots1 <- plots1 + new_plot
}
plots1 <- plots1 + plot_layout(ncol = 3)
for (i in 2:length(params2)) {
  new_plot <- create_traceplot(out, params2[i], names_df)
  plots2 <- plots2 + new_plot
}
plots2 <- plots2 + plot_layout(ncol = 3)
for (i in 2:length(params3)) {
  new_plot <- create_traceplot(out, params3[i], names_df)
  plots3 <- plots3 + new_plot
}
plots3 <- plots3 + plot_layout(ncol = 3)
```

```{r, echo = FALSE}
# create posterior summary dataframe
out_df <- rbind(out[[1]], out[[2]], out[[3]], out[[4]])

# get mean, sd, Rhat, effective sample size
posterior <- MCMCsummary(out)[, c("mean", "sd", "Rhat", "n.eff")]
posterior$short_names <- rownames(posterior)

# subset
posterior <- posterior[posterior$short_names %in% short_names, ]

# get 95% credibility interval
posterior$lower_ci <- NA
posterior$upper_ci <- NA
for (i in seq_len(nrow(posterior))) {
  sub <- out_df[, which(colnames(out_df) == posterior[i, "short_names"])]
  posterior[i, "lower_ci"] <- as.numeric(hdi(sub, 0.95)[2])
  posterior[i, "upper_ci"] <- as.numeric(hdi(sub, 0.95)[3])
}

# round
for (i in seq_len(nrow(posterior))) {
  
  dig <- 3 - floor(log10(abs(posterior[i, "mean"])))
  
  posterior[i, "mean"] <- round(posterior[i, "mean"], digits = dig)
  posterior[i, "sd"] <- round(posterior[i, "sd"], digits = dig)
  posterior[i, "lower_ci"] <- round(posterior[i, "lower_ci"], digits = dig)
  posterior[i, "upper_ci"] <- round(posterior[i, "upper_ci"], digits = dig)
}

posterior$mean <- as.character(posterior$mean)
posterior$sd <- as.character(posterior$sd)
posterior$lower_ci <- as.character(posterior$lower_ci)
posterior$upper_ci <- as.character(posterior$upper_ci)

# add real names
posterior <- left_join(names_df, posterior, by = "short_names")

# function for getting 95% credibility interval
get_ci <- function(lower, upper) {
  out <- paste0("(", lower, ", ", upper, ")")
  return(out)
}

# get final df
posterior_final <- posterior[, c("latex_names", "mean", "sd")]
posterior_final[, "95 CI"] <- mapply(get_ci,
                                     posterior$lower_ci,
                                     posterior$upper_ci)
posterior_final[, "Rhat"] <- posterior$Rhat
posterior_final[, "ESS"] <- posterior$n.eff
colnames(posterior_final) <- c("parameter", "mean", "sd",
                               "95 CI", "Rhat", "ESS")
```

**Table A5:** Posterior summaries, including the mean, standard deviation, 95% credibility interval (highest density interval), $\hat{R}$, and effective sample size. Descriptions of parameters can be found in Table 1. More information about parameter $\rho$ can be found in Appendix 2.1.

```{r, echo = FALSE}
# Process the table for parsed text
posterior_final <- posterior_final %>%
  mutate(parameter = paste0("$", parameter, "$"))

# Render table
knitr::kable(
  posterior_final,
  escape = FALSE,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE
) %>%
  kable_styling(full_width = FALSE, latex_options = c("repeat_header")) %>%
  kableExtra::row_spec(-1,
                       extra_latex_after = "\\renewcommand{\\arraystretch}{2}")
```
\newpage
**Figure A5:** Trace plots of posterior samples. Colors refer to separate chains. Descriptions of parameters can be found in Table 1.

```{r, echo = FALSE, fig.height=8.5}
plots1
```
\newpage

**Figure A5 continued.**

```{r, echo = FALSE, fig.height=8.5}
plots2
```

\newpage

**Figure A5 continued.**

```{r, echo = FALSE, fig.height=5.1}
plots3
```
