---
output: pdf_document
---

# Appendix 1

The integrated population model provides a framework for synthesizing multiple demographic datasets to mutually inform shared parameters. In Appendix 1.1, we further describe the Drayton Harbor time series data (D1). In Appendix 1.2, we describe additional information about the size-at-age data (D2), the model fitted to size-at-age data (D2), and posterior summaries. In Appendix 1.2, we describe the mark-recapture data (D3).

## Appendix 1.1

Below we describe the total removal effort and catch of European green crab (EGC, *C. maenas*) in Drayton Harbor (D1). While the removal count data is collected at many time points throughout the trapping season, we discretize the time intervals into biweekly periods in the model.

Drayton Harbor data were subset to only include recruits (age < 1) that appear in a single cohort in the fall of each year (Figure 1B). This subset only removed three recruit individuals in 2022 from the dataset, which  were from an off-cycle recruitment cohort.


```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)

# read in data
counts <- readRDS('../data/model_data/counts.rds')
f_index <- readRDS('../data/model_data/f_index.rds')
m_index <- readRDS('../data/model_data/m_index.rds')
s_index <- readRDS('../data/model_data/s_index.rds')
index_frac <- readRDS('../data/model_data/index_frac.rds')
totalt <- readRDS('../data/model_data/totalt.rds')

df_catch <- as.data.frame(matrix(NA, nrow=0,ncol=7))
colnames(df_catch) <- c('year','F (catch)','F (effort)',
                        'M (catch)','M (effort)','S (catch)','S (effort)')

years <- c('2020','2021','2022','2023')

for(i in 1:length(years)){
  for(t in 1:totalt[i]){
    count_f <- sum(rowSums(counts[t,,i,])*f_index[t,,i])
    count_s <- sum(rowSums(counts[t,,i,])*s_index[t,,i])
    count_m <- sum(rowSums(counts[t,,i,])*m_index[t,,i])
    effort_f <- sum(f_index[t,,i])
    effort_s <- sum(s_index[t,,i])
    effort_m <- sum(m_index[t,,i])
    
    vec <- c(years[i],count_f,effort_f,count_m,effort_m,count_s,effort_s)
    
    df_catch <- rbind(df_catch, vec)
  }
}

julian_day_df <- index_frac*365+60
julian_day <- c(julian_day_df[1:totalt[1],1],julian_day_df[1:totalt[2],2],
                julian_day_df[1:totalt[3],3],julian_day_df[1:totalt[4],4])
julian_day <- julian_day[!is.na(julian_day)]

df_catch <- cbind(julian_day, df_catch)

colnames(df_catch) <- c('Julian day','year','F (catch)','F (effort)',
                        'M (catch)','M (effort)','S (catch)','S (effort)')

```

**Table A1.1:** Number of crabs caught and total removal effort (number of traps) during each biweek and year. Julian day signifies the first Julian day of the discretized biweek period. F, M, and S correspond to Fukui, Minnow, and Shrimp traps, respectively.

```{r, echo = FALSE}
# Render table
knitr::kable(
  df_catch
) 
```


## Appendix 1.2

Below we provide additional information about the size-at-age data (D2), the model fitted to size-at-age data (D2), code used to fit data to the model, and posterior summaries.

### Appendix 1.2.1

Size-at-age data (D2) are collected from two sources: 1) records from crab removal observations in northeastern Pacific estuaries from Yamada et al. 2005, when the somatic growth of a strong recruitment class was tracked over time, and 2) crabs with an assigned year class from Drayton Harbor that were easily identifiable based on size and carapace color. While data from these Drayton Harbor recruits enter the integrated population model likelihood twice (D1 and D2), extensive simulation-based research has revealed that IPMs are robust to dependent data (Abadi 2010; McDonald 2016). Ages are assigned assuming recruitment settlement in March. **ADD CITATIONS**


```{r, echo = FALSE, warning = FALSE, message = FALSE}
growth_data <- read.csv('../data/growthdata.csv')

growth_summary <- growth_data %>% 
  group_by(Year,Waterbody) %>% 
  summarize(n=n())
```

**Table A1.2:** Number of size-at-age records during each year and waterbody in the northeastern Pacific.

```{r, echo = FALSE}
# Render table
knitr::kable(
  growth_summary
) 
```

### Appendix 1.2.2

The integrated population model (IPM) is fitted hierarchically, such that the seasonal growth parameters are fit with the size-at-age records, and the summarized posteriors are used to develop prior distributions in the overall IPM. Below, we detail the model used to fit the seasonal growth parameters with the size-at-age data.

The size-at-age data, $W_{a,i}$, indicating carapace width at age $a$ in year $i$ are fit to the same seasonal growth model as in the main text (Somers). Except here, we account for non-independence among data collected in the same year, since growth rate is likely affected by water temperature, which varies from year-to-year. 

The expected carapace width of a crab at age $a$ is:

$$
\widetilde{W}_{a,i} = y_{\infty}(1-exp(-k(a-t_0) - S(a) + S(t0))) + \epsilon_i
$$
$$
S(a) = \frac{Ck}{2\pi}sin(2\pi(a-t_s))
$$
$$
S(t_0) = \frac{Ck}{2\pi}sin(2\pi(t_0-t_s))
$$

To account for non-independence among data collected within the same year, we add a normally distributed error term with standard deviation term, $\tau_y$, to the equation for expected carapace width, $\widetilde{W}_{a,i}$:

$$
\epsilon_i \sim Normal(0, \tau_y)
$$

To account for individual variation in growth rate, the observed size-at-age data, $W_{a,i}$, follows a normal distribution, with the expected carapace width $\widetilde{W}_{a,i}$, and standard deviation, $\tau_{w}$.

$$
W_{a,i} \sim Normal(\widetilde{W}_{a,i}, \tau_{w})
$$

We use uninformative priors for the parameters:

$$
k \sim dunif(0,2)
$$
$$
C \sim Uniform(0,2)
$$
$$
t_s \sim Uniform(-1,0)
$$
$$
t_0 \sim Uniform(-1,1)
$$
$$
\tau_y \sim Uniform(0,100)
$$
$$
\tau_{w} \sim Uniform(0,100)
$$
$$
y_{\infty} \sim Uniform(80,140)
$$

### Appendix 1.2.3

Here we provide posterior summaries of the seasonal growth parameters fit with the size-at-age data. 

```{r, echo = FALSE}
library(MCMCvis)
library(bayestestR)
library(patchwork)

# read in samples
samples <- readRDS('../data/posterior_samples/savedsamples_growth.rds')
lower <- 100
upper <- 10001
```

```{r, echo = FALSE}
# create df with param names
short_names <- c("C","k","size_inf","t0","tau.ranef","tau.y","ts")
real_names <- c('C', 'k', 'y[infinity]', 't[0]', 'tau[w]','tau[y]','t[s]')
latex_names <- c('C', 'k', 'y_{\\infty}', 't_0', '\\tau_w','\\tau_y','t_s')

names_df <- as.data.frame(cbind(short_names,real_names,latex_names))
colnames(names_df) <- c('short_names','real_names','latex_names')
```

```{r, echo=FALSE}
# create posterior summary dataframe
out_sub <- list(samples[[1]][lower:upper,],samples[[2]][lower:upper,],
                samples[[3]][lower:upper,],samples[[4]][lower:upper,])
out_df <- rbind(samples[[1]][lower:upper,],samples[[2]][lower:upper,],
                samples[[3]][lower:upper,],samples[[4]][lower:upper,])

# get mean, sd, Rhat, effective sample size
posterior <- MCMCsummary(out_sub)[,c('mean','sd','Rhat','n.eff')]
posterior$short_names <- rownames(posterior)

# subset
posterior <- posterior[posterior$short_names %in% short_names,]

# get 95% credibility interval
posterior$lower_ci <- NA
posterior$upper_ci <- NA

for(i in 1:dim(posterior)[1]){
  sub <- out_df[,which(colnames(out_df)==posterior[i,'short_names'])]
  posterior[i,'lower_ci'] <- as.numeric(hdi(sub,0.95)[2])
  posterior[i,'upper_ci'] <- as.numeric(hdi(sub,0.95)[3])
}

# round
for(i in 1:dim(posterior)[1]){
  if(posterior[i,'short_names'] == 'size_inf'){
    posterior[i,'mean'] <- round(posterior[i,'mean'], digits = 2)
    posterior[i,'sd'] <- round(posterior[i,'sd'], digits = 2)
    posterior[i,'lower_ci'] <- round(posterior[i,'lower_ci'], digits = 2)
    posterior[i,'upper_ci'] <- round(posterior[i,'upper_ci'], digits = 2)
  } else {
    posterior[i,'mean'] <- round(posterior[i,'mean'], digits = 3)
    posterior[i,'sd'] <- round(posterior[i,'sd'], digits = 3)
    posterior[i,'lower_ci'] <- round(posterior[i,'lower_ci'], digits = 3)
    posterior[i,'upper_ci'] <- round(posterior[i,'upper_ci'], digits = 3)
  }
}
posterior$mean <- as.character(posterior$mean)
posterior$sd <- as.character(posterior$sd)

# add real names
posterior <- left_join(names_df, posterior, by = 'short_names')

# function for getting 95% credibility interval 
get_ci <- function(lower, upper){
  out <- paste0('(',lower, ', ', upper, ')')
  return(out)
}

# get final df
posterior_final <- posterior[,c('latex_names','mean','sd')]
posterior_final[,'95 CI'] <- mapply(get_ci, 
                                     posterior$lower_ci, 
                                     posterior$upper_ci)
posterior_final[,'Rhat'] <- posterior$Rhat
posterior_final[,'ESS'] <- posterior$n.eff
posterior_final[,'IPM prior'] <- c('X','X','X','','','','X')
colnames(posterior_final) <- c('parameter','mean','sd',
                               '95 CI','Rhat','ESS','IPM prior')

```

```{r, echo=FALSE}
# Process the table for parsed text
posterior_final <- posterior_final %>%
  mutate(
    parameter = paste0("$", parameter, "$")
  )

# Render table
knitr::kable(
  posterior_final,
  caption = "Table A1.2: Posterior summaries, including the mean, standard deviation, 95\\% credibility interval (highest density interval), Rhat, and effective sample size. An X in 'IPM prior' indicates that the posterior is used in the overall IPM. Descriptions of parameters can be found in Appendix 1.2.2.",
  escape = FALSE,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE
) %>%
  kable_styling(full_width = FALSE, latex_options = c("repeat_header")) %>%
  kableExtra::row_spec(0, extra_latex = "\\renewcommand{\\arraystretch}{1.5}")
```


```{r, echo = FALSE, results = 'hide'}
# create functions
create_traceplot <- function(out,lower,upper,param,names_df){
  name <- names_df$real_names[which(names_df$short_names==param)]
  
  plot <- ggplot()+
  geom_line(aes(x=lower:upper,
                y=out[[1]][lower:upper,param]),
            color='pink')+
  geom_line(aes(x=lower:upper,
                y=out[[2]][lower:upper,param]),
            color='purple')+
  geom_line(aes(x=lower:upper,
                y=out[[3]][lower:upper,param]),
            color='blue')+
  geom_line(aes(x=lower:upper,
                y=out[[4]][lower:upper,param]),
            color='red')+
    scale_x_continuous(breaks=c(2000,6000,10000))+
  labs(x='iteration',y='value')+
    ggtitle(parse(text=name))+
    theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
  return(plot)
}
```

```{r, echo = FALSE, results = 'hide'}
# create trace plots

plots <- create_traceplot(samples, lower, upper, 
                          names_df$short_names[1], names_df)

for(i in 2:length(names_df$short_names)){
  new_plot <- create_traceplot(samples, lower, upper, 
                               names_df$short_names[i], names_df)
  plots <- plots + new_plot
}
plots <- plots + plot_layout(ncol=3)
```

\newpage

**Figure A1:** Trace plots of posterior samples of size-at-age model. Colors refer to separate chains. Descriptions of parameters can be found in Appendix 1.2.2.

```{r, echo = FALSE}
plots
```


