---
output: pdf_document
---
# Supplemental Information

**Figure S1:** Population forecasts in response to varying removal efforts. Size distributions show the equilibrium crab abundance in each size class, $N(y)^E$, at the end of the year after overwinter mortality when *A.* 0 traps, *B.* 28 traps, *C.* 112 traps, *D.* 560 traps, *E.* 840 traps, *F.* 1400 traps, and *G.* 2800 traps were applied evenly over a trapping season of 14 biweeks. Solid line indicates the median size-structured abundance across simulation replicates, and the shaded area indicates $\pm1$ standard deviation across simulation replicates. Colors indicate trap type used (i.e., in panel B, the purple line shows the resulting size distribution after a trapping effort of 28 Minnow traps).
```{r, echo=FALSE, out.width = "100%"}
knitr::include_graphics(
  paste0("C:/Users/abiga/Documents/Berkeley/Structured_Decision_Making/",
         "pop_dynamics_model/IPMsquared/figures/supplemental_IPM_simulations_mean.png")
)
```

\newpage

**Figure S2:** Population forecasts in response to varying removal efforts, relative to no removal. Size distributions show the ratio of the equilibrium crab abundance in each size class, $N(y)^E$, when no removal occurred ($N(y)^E_{effort=0}$) relative to a removal effort greater than zero ($N(y)^E_{effort>0}$). Ratios are calculated based on size-structured abundance at the end of the year after overwinter mortality when *A.* 112 traps, *B.* 560 traps, and *C.* 2800 traps were applied evenly over a trapping season of 14 biweeks. A ratio of one means that the size-structured abundance after no removal equals the size-structured abundance after application of X traps. A ratio less than one means that the application of X traps removes decreases the size-structured abundance, relative to no removal. A ratio greater than one means that the application of X traps increases the size-structured abundance, relative to no removal. Solid line indicates the median size-structured abundance across simulation replicates, and the shaded area indicates $\pm1$ standard deviation across simulation replicates. Colors indicate trap type used (i.e., in panel B, the purple line shows the resulting size distribution ratio after a trapping effort of 0 traps, relative to 560 Minnow traps).
```{r, echo=FALSE}
knitr::include_graphics(
  paste0("C:/Users/abiga/Documents/Berkeley/Structured_Decision_Making/",
         "pop_dynamics_model/IPMsquared/figures/supplemental_IPM_simulations_diff.png")
)
```

\newpage

**Table S1**: Total equilibrium abundance across all size classes, $N_{total}^E$ for population forecasts in response to varying removal efforts. Mean refers to the mean $N_{total}^E$ across all simulation replicates, and sd refers to the standard deviation of $N_{total}^E$ across all simulation replicates.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(kableExtra)

# read in simulation results
out <- readRDS("C:/Users/abiga/Documents/Berkeley/Structured_Decision_Making/pop_dynamics_model/IPMsquared/data/simulations/simulations_effort.rds")
out_noeffort <- readRDS("C:/Users/abiga/Documents/Berkeley/Structured_Decision_Making/pop_dynamics_model/IPMsquared/data/simulations/simulations_noeffort.rds")

# summarize
summary <- as.data.frame(matrix(NA, nrow = 19, ncol = 4))
colnames(summary) <- c("type", "effort", "$N_{total}^E$ (mean)", "$N_{total}^E$ (sd)")

# no effort
summary[1, ] <- c("NA", 0, round(mean(rowSums(out_noeffort)), 1), 
                  round(sd(rowSums(out_noeffort)), 1))

# effort summaries
summary[2:19, ] <- as.data.frame(out) %>% 
  mutate(N = rowSums(across(1:22))) %>% 
  mutate(type = case_when(prop_s == 1 ~ "Shrimp",
                          prop_f == 1 ~ "Fukui",
                          prop_m == 1 ~ "Minnow")) %>% 
  group_by(type, nobs) %>% 
  summarise(mean = round(mean(N), 1), sd = round(sd(N), 1))
```

```{r, echo = FALSE}
# Render table
knitr::kable(
  summary,
  escape = FALSE,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE
) %>%
  kable_styling(full_width = FALSE, latex_options = c("repeat_header")) %>%
  kableExtra::row_spec(-1,
                       extra_latex_after = "\\renewcommand{\\arraystretch}{2}")
```
