\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{float}
\usepackage{longtable}
\usepackage{caption}
\usepackage{setspace} 
\usepackage{lineno}
\usepackage{authblk}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{url}
\usepackage{hyperref}
\usepackage[
backend=biber,
style=apa,
sorting=nyt
]{biblatex}
\addbibresource{maintext.bib}

\title{An integrated integral projection model (IPM\textsuperscript{2}) to disentangle size-structured harvest and natural mortality}
\author[1,*]{Abigail G. Keller}
\author[2]{Benjamin R. Goldstein}
\author[3]{Leah Skare}
\author[1]{Perry de Valpine}
\affil[1]{\small Department of Environment Science, Policy, and Management, University of California, Berkeley, Berkeley, California, USA}
\affil[2]{\small Department of Forestry and Environmental Resources, North Carolina State University, Raleigh, NC, USA}
\affil[3]{\small Northwest Straits Commission, Washington Department of Ecology, Mount Vernon, WA, USA}
\affil[*]{\small Corresponding author: Abigail G. Keller, agkeller@berkeley.edu}
\date{}

\begin{document}

\doublespacing

\linenumbers

\maketitle

\section{Acknowledgements}

We thank Sylvia Yamada, Ted Grosholz, Brian Turner, Tom Therriault, and Brett Howard for providing valuable feedback on European green crab population dynamics and model development; Northwest Straits Commission, Washington Department of Fish and Wildlife, community volunteers, Veterans Conservation Corps, Washington Conservation Corps, and Washington Sea Grant for data collection and management. Mark-recapture data from Roche Cove was collected as part of a project (Towards Absolute Population Estimates for European Green Crab Management) funded by Fisheries and Oceans Canada’s Aquatic Invasive Species Science program to Tom Therriault with field support from Hailey Shchepanik and Karen Dyke.
This material is based upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research, under Award Number DE-SC0024386. This material is based in part upon work supported by the National Science Foundation under Grant No. DBI-1942280 and a grant from U.S. Geological Survey.


\section{Disclaimer}

This report was prepared as an account of work sponsored by an agency of the United States Government. Neither the United States Government nor any agency thereof, nor any of their employees, makes any warranty, express or implied, or assumes any legal liability or responsibility for the accuracy, completeness, or usefulness of any information, apparatus, product, or process disclosed, or represents that its use would not infringe privately owned rights. Reference herein to any specific commercial product, process, or service by trade name, trademark, manufacturer, or otherwise does not necessarily constitute or imply its endorsement, recommendation, or favoring by the United States Government or any agency thereof. The views and opinions of authors expressed herein do not necessarily state or reflect those of the United States Government or any agency thereof.

\section{Conflict of interest}

The authors declare no conflict of interest.

\section{Author contributions}

A.G. Keller, B.G., Goldstein, and P. de Valpine conceived the ideas and designed methodology; L. Skare collected the data; A.G. Keller and P. de Valpine analysed the data; A.G. Keller led the writing of the manuscript. All authors contributed critically to the drafts and gave final approval for publication.

\section{Data availability statement}

All data and R code associated with this analysis can be found in the Zenodo archive (\url{https://doi.org/10.5281/zenodo.16995754})
and GitHub (\url{https://github.com/abigailkeller/IPMsquared}). 

\end{document}